---
title: "Web Data Collection with R"
author: "Peter Meißner / 2016-02-29 --  2016-03-04 / ECPR WSMT"
header-includes:
  - \definecolor{links}{HTML}{800080}
  - \hypersetup{colorlinks,linkcolor=,urlcolor=links}
output:
  beamer_presentation:
    fonttheme: structurebold
    keep_tex: yes
    toc: yes
  slidy_presentation: default
subtitle: File Manipulation
keep_tex: yes
---

```{r, include=FALSE}
options(width = 40)
```


# Why ? 

# Functions

## Patterns

function      | description
------------- | -------------
`list.files()`        | list files 
`list.dirs()`         | list directories 
`file.info()`         | information on file
`file.size()`         | check size of file 
`file.exists()`       | check existence of file
`file.copy()`         | copy file
`file.remove()`       | delete file 
`file.rename()`       | rename file
`file.mtime()`        | get modiefied time of file 
`readLines()`         | read file into char. vector
`writeLines()`        | write character vector into file
`write()`             | write to file, e.g. append to file 

&nbsp;
    
## use case 1 - write file if not existent
```{r, include=FALSE}
file.remove("testfile.txt")
```

```{r}
fname <- "testfile.txt"

if( !file.exists(fname)  ) {
  writeLines( as.character(Sys.time()), fname )
}

file.mtime(fname)
readLines(fname)
```

## use case 2 - append to file 

```{r}

for( i in 1:3 ){
  atime <- substr(as.character(Sys.time()),12,20)
  mtime <- substr(as.character(file.mtime(fname)),12,20)
  write(
    paste0(
      " append : ", atime, "   mtime : ", mtime, 
      collapse = ""
    ), 
    fname, 
    append = TRUE
  )
  Sys.sleep(1)
}
```

## use case 2 - append to file

```{r}
file.mtime(fname)
readLines(fname)
```


## use case 3 - get file name from URL

```{r}
url      <- "http://www.r-datacollection.com/materials/http/abunchofasciis/index.html"
destfile <- basename(url)
download.file(url = url, destfile = destfile)
```

```{r}
destfile
file.exists(destfile)
readLines(destfile, warn=FALSE)
```


## use case 4 - handling files names in a loop

```{r}
for ( i in 1:10) {
  content <- paste(sample(letters,10),collapse = "")
  fname   <- 
    paste(
      "f_", 
      Sys.time(), " ", 
      stringr::str_pad(i, 3, "left", 0),  
      ".test",
      collapse = "", sep=""
    )
  writeLines(content, fname)
}
```

## use case 4 - handling files names in a loop

```{r}
list.files(pattern = "\\d.test$")
```


## use case 5 - zipping 

```{r}
filelist <- list.files(pattern = ".txt$|html$|test$")
zip(zipfile = "archive.zip", files = filelist)
file.remove(filelist[file.exists(filelist)])
list.files()
```





    
    

