---
title: "Web Data Collection with R"
author: "Peter Meißner / 2016-02-29 --  2016-03-04 / ECPR WSMT"
header-includes:
  - \definecolor{links}{HTML}{800080}
  - \hypersetup{colorlinks,linkcolor=,urlcolor=links}
output:
  beamer_presentation:
    fonttheme: structurebold
    keep_tex: yes
    toc: yes
  slidy_presentation: default
subtitle: robots.txt
keep_tex: yes
---



## teaser

- robots.txt files are part of many webpages
- the idea is ask bots (e.g. googlebot, etc.) t adhere to those rules and not cause any trouble
- for us this is valuable information about what might be unwanted behaviour as well

## exmaple and syntax : https://www.researchgate.net/robots.txt

```robotstxt
User-agent: *
Allow: /
Disallow: /connector/
Disallow: /deref/
Disallow: /plugins.
Disallow: /firststeps.
Disallow: /publicliterature.PublicLiterature.search.html

Sitemap: https://www.researchgate.net/sitemaps/sitemap-index.xml
```

## there is a package for that

```{r}
library(robotstxt)
??robotstxt

rtxt <- 
  robotstxt$new(
    domain="https://www.researchgate.net"
  )
```

## there is a package for that

```{r}
rtxt
```

## there is a package for that

```{r}
rtxt$check(c("/connector/index.html","index.html"))
```

## But always have a look at the ToS

https://www.researchgate.net/application.TermsAndConditions.html

ARTICLE 5: MISUSE OF THE SERVICE

- Users must not misuse the Service. 
- Misuse of the Service includes, without limitation: 
    - insults to other Users;
    - automated or massive manual retrieval of other Users' profile data ("data harvesting");
    - advertising for commercial products or services of all kinds;
    - unsolicited job offers and business proposals;
    - all kinds of technical attacks on the servers.
- All aforementioned behaviors in this article are strictly forbidden, unless the User has obtained prior written permission by the Provider.



