---
title: 'Exercise: RegEx'
author: "Peter Meißner"
date: "28 Februar 2016"
output:
  html_document:
    keep_md: no
    toc: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##1) Character Encoding

```{r}
readLines("text1.txt")
readLines("text2.txt")
```

a) use the encoding option of readLines() to read show files correctly

```{r}
readLines("text1.txt", encoding = "UTF-8")
readLines("text2.txt", encoding = "latin1")
```


b) use Encoding to show the files correctly

```{r}
text1 <- readLines("text1.txt")
Encoding(text1) <- "UTF-8"
text1

text2 <- readLines("text2.txt")
Encoding(text2) <- "latin1"
text2
```

c) use iconv to solve the problem 

```{r}
text1 <- readLines("text1.txt")
text1 <- iconv(text1, "UTF-8", "UTF-8")
text1

text2 <- readLines("text2.txt")
text2 <- iconv(text2, "latin1", "UTF-8")
text2
```

d) use stringi::stri_enc_detect() to guess Encoding of the files
```{r}
library(stringi)

text1 <- readLines("text1.txt")
stri_enc_detect(text1)
stri_enc_detect2(text1)

text2 <- readLines("text2.txt")
stri_enc_detect(text2)
stri_enc_detect2(text2)
```

e) write your name with `\u0000` expressions (http://unicode-table.com/en/), e.g.:

```{r}
name <- "\u0055\u0064\u006f"
name
```

f) read in 17814-0.txt and get the encoding right

```{r}
readLines("17814-0.txt", encoding = "UTF-8")[1:10]
```

##2) Information Extraction

a) read in pg345.txt
    - count how many times the word "blood" or "Blood" is used throughout the book
    - use grep() to get an index of lines containing the word
    - use hist(...,n=100000) to make a "zebra?"-chart
    - what might be other interesting words?
    - can you add them to your RegEx?

```{r}
library(stringr)
text <- readLines("pg345.txt")

sum(str_count(text, "[bB]lood"))
sum(str_count(text, "\\b[bB]lood\\b"))

grep("\\b[bB]lood\\b", text)

hist(grep("\\b[bB]lood\\b", text),n=100000, main="Blood in Dracula", xlab="line number")
box()

hist(grep("\\b[bB]lood\\b|\\b[fF]ear\\b", text),n=100000, main="Blood and Fear in Dracula", xlab="line number")
box()

```


b) read in pg345.txt
    - use `paste(..., collapse="\n")` to combine the text into one single string
    - use str_split() to split this string into words
    - use tabulate() and sort to get the 10 most frequent words as well as the 10 least frequent words
    
c) for all files: 17814-0.txt, pg11.txt, pg1661.txt, pg174.txt, pg2600.txt, pg345.txt, pg34901.txt    
    - find a way to extract:
        - title
        - author
        - posting date
    - find a way to drop information added by Projct Gutenberg

```{r}
txt <- readLines("pg11.txt")
txt[1:10]

grep("^Title", txt, value = TRUE)
textfiles <- c("17814-0.txt", "pg11.txt", "pg1661.txt", "pg174.txt", "pg2600.txt", "pg345.txt")
TXT <- list()
for( i in seq_along(textfiles) ){
  TXT[[i]] <- readLines(textfiles[i])
}

grep("^Title", TXT[[1]], value = TRUE)[1]

get_gutenberg_title <- function(fname){
  txt <- readLines(fname)
  tmp <- grep("^Title", txt, value = TRUE)[1]  
  str_replace(tmp, "^Title: ", "")
}

get_gutenberg_author <- function(fname){
  txt <- readLines(fname)
  tmp <- grep("by", txt, value = TRUE)[1]
  str_replace(tmp, "^.*by ","")
}

get_gutenberg_posting_date <- function(fname){
  txt <- readLines(fname)
  tmp <- grep("posting date|release date", txt, value = TRUE, ignore.case = TRUE)[1]
  str_replace_all(tmp, "^.*: | \\[.*$","")
}

lapply(textfiles, get_gutenberg_title)
lapply(textfiles, get_gutenberg_author)
lapply(textfiles, get_gutenberg_posting_date)
```



d) read in 2012.txt and build a data.frame containing the following information
    - name
    - number of reviews
    - institution

```{r}
txt <- readLines("2012.txt", warn = FALSE)[-c(1:52)]
txt <- paste0(txt, collapse="") 

txt <-
  str_split(txt, "\\)") %>% 
  unlist()

txt <- 
  txt %>% 
  str_replace_all("\t|  \f"," ") %>% 
  str_trim() %>% 
  str_replace_all("  ","") 

name <- 
  txt %>% 
  str_extract("^.*\\.") %>% 
  str_replace("\\.","")

reviews <- 
  txt %>% 
  str_extract("\\d") %>% 
  as.numeric()

institution <- 
  txt %>% 
  str_extract("\\..*\\(") %>% 
  str_replace_all("\\.|\\(","") %>% 
  str_trim()
```















