---
title: "Exercise: RegEx"
author: "Peter Meißner"
date: "28 Februar 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##1) Character Encoding

```{r}
readLines("text1.txt")
readLines("text2.txt")
```

a) use the encoding option of readLines() to read show files correctly

b) use Encoding to show the files correctly

c) use iconv to solve the problem 

d) use stringi::stri_enc_detect() to guess Encoding of the files

e) write your name with `\u0000` expressions (http://unicode-table.com/en/), e.g.:

```{r}
name <- "\u0055\u0064\u006f"
name
```

f) read in 17814-0.txt and get the encoding right


##2) Information Extraction

a) read in pg345.txt
    - count how many times the word "blood" or "Blood" is used throughout the book
    - use grep() to get an index of lines containing the word
    - use hist(...,n=100000) to make a "zebra?"-chart
    - what might be other interesting words?
    - can you add them to your RegEx?

b) read in pg345.txt
    - use paste(..., collapse="\n") to combine the text into one single string
    - use str_split() to split this string into words
    - use tabulate() and sort to get the 10 most frequent words as well as the 10 least frequent words
    
c) for all files: 17814-0.txt, pg11.txt, pg1661.txt, pg174.txt, pg2600.txt, pg345.txt, pg34901.txt    
    - find a way to extract:
        - title
        - author
        - posting date
    - find a way to drop information added by Projct Gutenberg

d) read in 2012.txt and build a data.frame containing the following information
    - first name
    - last name
    - number of reviews
    - institution
















