---
title: 'Exercise: Xpath'
author: "Peter Meißner"
date: "29 February 2016"
output:
  html_document:
    keep_md: no
    toc: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##1) xml2 package

a) getting started
- load the xml2 package
- use `ls(package:xml2)` to get a list of availible functions
- use `read_xml()` to read in the following resource: http://www.r-datacollection.com/materials/ch-4-xpath/potus/potus.xml and save it myxml

```{r}
library(magrittr)
library(xml2)
ls(package:xml2)

if( !file.exists("myxml.xml") ){
  url <- "http://www.r-datacollection.com/materials/ch-4-xpath/potus/potus.xml"
  download.file(url, "myxml.xml")
}

xml <- read_xml("myxml.xml")
```

- use your browser to look at the XML

```{r, eval=FALSE}
browseURL(url)
```


b) convenience functions
- how many entries has the XML, which function did you use

```{r}
xml_length(xml)
```

- extract all the text of the XML and use cat() to print the text to the screen

```{r}
cat(xml_text(xml))
```

- use a function of the xml2 package to have a look at the structure of the XML

```{r}
xml_structure(xml)
```


c) extraction

- use XPath to get all religion nodes

```{r}
xml_find_all(xml, xpath = "//religion")
```

- use CSS-Selectors to get all religion nodes

```{r}
religion <- xml_find_all(xml, "religion")
religion
```

- extract the text from the nodes and use table() and sort() to produce a sorted table

```{r}
religion %>% 
  xml_text() %>% 
  table() %>% 
  sort(TRUE) %>% 
  as.data.frame()
```

- use xml_find() in combination with XPath to extract the node for the first president

```{r}
xml %>% 
  xml_find_one(xpath="//president[1]")
```

- extract the name node for all presidents after the 40th

```{r}
xml %>% 
  xml_find_all(xpath="//president[position()>40]")
```

- using XPath get the nodes for Pierce, Garfield and Taylor


```{r}
xml %>% 
  xml_find_all(xpath="//president[contains(./name, 'Pierce') or contains(./name, 'Garfield') or contains(./name, 'Taylor')]")
```

- extract the occupation

```{r}
xml %>% 
  xml_find_all("occupation") %>% 
  xml_text() %>% 
  table() %>% 
  sort(TRUE) %>% 
  as.data.frame()
```

- transform the XML into a data.frame

```{r}
tagnames <- 
  xml %>% 
  xml_find_all(xpath="/*/*/*") %>% 
  xml_name() %>% 
  unique()

df <- data.frame(NA)
for( tag in tagnames ){
    tmp <- 
      xml %>% 
      xml_find_all(paste0("//",tag)) %>%
      xml_text()
  df <- data.frame(df, tmp)
}
df <- df[, -1]
names(df) <- tagnames

df

```


























