---
title: "Exercise: file manipulation"
author: "Peter Meißner"
date: "28 Februar 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1) Logging the System time

a) write a loop that 
    - sleeps 1 second
    - writes a text-file with name equal to the current time
    
b) write a function that 
    - writes a file with some content of your choosing
    - but only if a file with the same name does not exist or is older than 20 seconds
    
*example of a function writing a file*
```{r}
  write_punny_file <- function(name="no_a_clever_name.txt"){
    writeLines("content", name)
  }
```

c) write a loop that 
    - sleeps for a second
    - appends the current time to a file


##2) Pasting file names
    
a) generate a data.frame 
    - with two variables 
    - each having three values 
    - build a double loop (one over variable 1, one over variable 2) 
    - for each iteration a file should be written to disk 
    - the file name should entail the value of variable one and two
  
*example of a double loop*
```{r, eval=FALSE}
  for( i in 1:3){
    for( j in 1:3){
      cat("- i=", i, "j=", j, "\n")
    }
  }
```
  
  
b) use expand.grid() on the data.frame created before
    - than use apply() to apply the paste function with option collapse="_" on each row
    - use the resulting strings to create the same file as in ex-2.a



##3) Reading files
  
a) use list.files() to get a vector of file names in the current folder

b) use list.files() in combination with option pattern 
    - to only get a subset of files
    - use three different subsets
    
c) make an empy list
    - use list.files() to get a vector of file names
    - use file.info() to get further information on the files (mtime)
    - read in the content of the files 
        - only for those files that are newest according to mtime
  
  
  
  
  
        
