---
title: 'Exercise: file manipulation'
author: "Peter Meißner"
date: "28 Februar 2016"
output:
  html_document:
    keep_md: no
    toc: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(magrittr)
library(stringr)
```


```{r cleaning up}
if( grepl("file_manipulation_exercise",basename(getwd())) ){
  list.files(pattern = "txt$")  %>% 
    file.remove()
  
  list.files(pattern = "^\\d.*\\d$") %>% 
    file.remove()
  
  file.remove("mylogfile")
}

```


##1) Logging the System time

a) write a loop that 
    - sleeps 1 second
    - writes a text-file with name equal to the current time

```{r}
for(i in 1:10){
  Sys.sleep(1)
  
  Sys.time() %>% 
    as.character() %>% 
    str_replace_all(":", "_") %>% 
    writeLines( "blah", .)
}

list.files(pattern = "^\\d")
```

    
b) write a function that 
    - writes a file with some content of your choosing
    - but only if a file with the same name does not exist or is older than 20 seconds
    
*example of a function writing a file*
```{r}
  write_punny_file <- function(name="not_a_clever_name.txt"){
    timediff <-
      as.numeric(Sys.time()) -
    as.numeric(file.info(name)$mtime)
    
    if( !file.exists(name) | timediff > 20 | is.na(timediff > 20) ){
      writeLines("content", name)
      return(TRUE)
    }else{
      return(FALSE)
    }
  }

write_punny_file()
list.files(pattern = "^not_")

Sys.time()
file.info("not_a_clever_name.txt")$mtime
```

c) write a loop that 
    - sleeps for a second
    - appends the current time to a file
```{r}
for( i in 1:10){
  write(
    as.character(Sys.time()), 
    "mylogfile", append = TRUE)  
  Sys.sleep(0.3)
}

readLines("mylogfile")
```


##2) Pasting file names
    
a) generate a data.frame 
    - with two variables 
    - each having three values 
    - build a double loop (one over variable 1, one over variable 2) 
    - for each iteration a file should be written to disk 
    - the file name should entail the value of variable one and two
  
*example of a double loop*
```{r, eval=FALSE}
  for( i in 1:3){
    for( j in 1:3){
      cat("- i=", i, "j=", j, "\n")
    }
  }
```

*solution*

```{r}
var1 <- letters[1:3]
var2 <- LETTERS[7:9]
df  <- data.frame(var1, var2)

  for( i in seq_along(var1)){
    for( j in seq_along(var2)){
      txt   <- paste("var1 =", df$var1[i], "\nvar2 =", df$var2[j])
      fname <- paste0(df$var1[i], "_", df$var2[j], ".txt")
      writeLines(txt, fname)
    }
  }

list.files(pattern = "txt$")

readLines("a_G.txt")
readLines("c_I.txt")
```
  
b) use expand.grid() on the data.frame created before
    - than use apply() to apply the paste function with option collapse="_" on each row
    - use the resulting strings to create the same file as in ex-2.a

```{r}
df_expanded <- expand.grid(df)
df_expanded

df_expanded$fnames <-
  df_expanded %>% 
  apply(1, paste0, collapse="_") %>% 
  paste0(".txt")

for( i in seq_along(df_expanded[,1]) ){
  with(
    df_expanded,
    writeLines(
      paste(".var1 =", var1[i], "\n.var2 =", var2[i]),
      fnames[i],
    )
  )
}

list.files(pattern = "txt$")

readLines("a_G.txt")
readLines("c_I.txt")
```



##3) Reading files
  
a) use list.files() to get a vector of file names in the current folder

```{r}
list.files()
```


b) use list.files() in combination with option pattern 
    - to only get a subset of files
    - use three different subsets
    
```{r}
list.files(pattern = "^\\d")

list.files(pattern = "txt$")

list.files(pattern = "manipulation")

list.files(pattern = "manipulation", full.names = TRUE, recursive = TRUE)
```
    
c) make an empy list
    - use list.files() to get a vector of file names
    - use file.info() to get further information on the files (mtime)
    - read in the content of the files 
        - only for those files that are newest according to mtime
  
  
```{r}
fnames <- list.files(full.names = TRUE, recursive = TRUE)

info   <- file.info(fnames)
mtime  <- info$mtime

newest <- fnames[order(mtime, decreasing = TRUE)]

texts  <- lapply(newest, readLines, warn=FALSE)
```
