---
title: "Web Data Collection with R"
author: "Peter Meißner / 2016-02-29 --  2016-03-04 / ECPR WSMT"
header-includes:
  - \definecolor{links}{HTML}{800080}
  - \hypersetup{colorlinks,linkcolor=,urlcolor=links}
output:
  beamer_presentation:
    fonttheme: structurebold
    keep_tex: yes
    toc: yes
  slidy_presentation: default
subtitle: Character Encoding
keep_tex: yes
---



# Character Encodings

##Character Encodings

Character Encodings are ... 

- are like family ...
- ... some of them you do not like but cannot avoid ...
- ... something we will struggle with but have cope anyways

The best thing is ...

- R has them all 

The worst thing is ... 

- R has them all 

##Character Encodings
- computers store everything as 0s and 1s (bits)
- in cs there are differing layers of abstraction
- one bit of information is called bit
- bits are quite uninformative as they only ave two states
- so they are are grouped into bytes (8 bits)
- one byte can have 256 different values (2^8)
- so it can store numbers 0 to 255 or 1 to 256 or ... -127 to 128
- or it can map to characters e.g. ASCII (abcABC.:-_,;#'+*~|<>!"§$%&/()=?}][{\}^°, ...")
- ASCII is a character set - the set of characters you want to be able to store - even 7 Bits would suffice to store it

##Character Encodings
- for larger character sets than ASCII (ä ö ü é è ... ) on needs to get clever since one byte does not suffice to map all characters to 0s and 1s
- unfortunate people got clever in differing ways 
    1) using more than one byte to map more characters ('wide' characters, UTF-16, USC-2, Windows OSs)
    2) using one or more bytes and using the first byte to encode how manies are used ('multi-byte characters', UTF-8, Unix based OSs)
- otherwise we would not have to talk about character sets and character encodings
    

##Character Encodings
```{r}
rawToBits(as.raw(62:66)) # as bits
as.raw(62:66) # bytes as hexa-decimal
as.numeric(as.raw(62:66)) # as numbers
rawToChar(as.raw(62:66)) # bytes as chararcters
```


##A character set problem
```{r}
text           <- rawToChar(as.raw(228))
Encoding(text) <- "UTF-8" 
text
Encoding(text) <- "latin1" 
text
```

Results differ because for latin1 character 228 is know but not for UTF-8

##An encoding problem
Of cause UTF-8 knows how to encode "ä" ... 
```{r}
text <- "ä"
charToRaw(text)
Encoding(text) <- "latin1"
text
```
... but here the results differ because "UTF-8" has another system translating characters to bytes. In latin1 the two bytes are interpreted as two characters. 



##Which default encoding does your R use
```{r}
Sys.getlocale()

# if yor locale is something other than UTF-8, 
# switch 'latin1' and 'UTF-8' and you shall be good to go
```

##Changing interpretation of bytes
```{r}
text <- "Små grodorna, små grodorna är lustiga att se."
Encoding(text) <- "UTF-8"
text
```

##Changing interpretation of bytes
```{r}
text <- "Små grodorna, små grodorna är lustiga att se."
Encoding(text) <- "latin1"
text
```

##Changing bytes and interpretation
```{r}
text <- "Små grodorna, små grodorna är lustiga att se."
text <- iconv(text, "UTF-8", "latin1")
Encoding(text)
text
```

##Noe that all sources might have another encoding than your R default locale!
```{r}
text <- "Små grodorna, små grodorna är lustiga att se."
text <- iconv(text, "UTF-8", "latin1")
writeLines(text, "text_latin1.txt", useBytes = TRUE)
text <- readLines("text_latin1.txt")
Encoding(text)
text
Encoding(text) <- "latin1"
text
```























