---
title: "Web Data Collection with R"
author: "Peter Meißner / 2016-02-29 --  2016-03-04 / ECPR WSMT"
header-includes:
  - \definecolor{links}{HTML}{800080}
  - \hypersetup{colorlinks,linkcolor=,urlcolor=links}
output:
  beamer_presentation:
    fonttheme: structurebold
    keep_tex: yes
    toc: yes
  slidy_presentation: default
subtitle: Regular Expressions / RegEx - A Case Study
keep_tex: yes
---


# Getting to know the page

## first glance at: http://ajps.org/list-of-reviewers/
```{r, eval=FALSE}
url <- "http://ajps.org/list-of-reviewers/"
browseURL(url)
```

## getting to know the page
- look at the source code (Cntr-U)
- inspecting elements (Cntr-Shift-I)

## surprise
- reviewer lists are not part of the web page but available as PDF downloads



# Scraping Strategy

## getting PDFs
1) download page / load into R 
    - `read_html()` *[rvest]*
2) extract anker nodes `<a ...>` 
    - `html_nodes(..., xpath=...)` *[rvest]*  
3) extract `href` attribute from nodes
    - `html_attr(..., "href")` *[rvest]*
4) filter links (keep those entailing: 'review'; four digits; 'pdf')
    - `str_detect(..., "review.*\\d{4}.*pdf")` *[stringr]*
5) download PDFs to disk
    - `download.file(..., ..., mode="wb")` *[utils]*

## extracting information from PDF
6) converting PDF to something we can work with
    - e.g. Adobe Acrobat Pro
        - HTML, XML, TXT, ...   
    - e.g. Xpdf (http://www.foolabs.com/xpdf/download.html)
        - HTML, TXT, ... 
        - WINDOWS: http://www.foolabs.com/xpdf/download.html - add install path to path variable / see: http://www.computerhope.com/issues/ch000549.htm
        -  Linux e.g.: `sudo apt-get install poppler-utils`
7) load into R and use Regular Expressions extract information




# Scraping 




## getting PDFs
```{r, message=FALSE}
# packages needed
require(rvest)
require(stringr)
```

## getting PDFs
```{r, cache=TRUE}
# url with list of reviews
url <- "http://ajps.org/list-of-reviewers/"

# get page
content <- read_html(url)

# get anker (<a href=...>) nodes via xpath
ankers  <- html_nodes(content, xpath = "//a")

# get value of ankers' href attribute
hrefs   <- html_attr(ankers, "href", 
                     default="NO HREF IN HERE")
```


## getting PDFs
```{r, cache=TRUE}
# filter links: should entail ... 
# 'review', four-digit number, 'pdf'
pdf <- hrefs[ str_detect(hrefs, "review.*\\d{4}.*pdf") ]
pdf
```


## getting PDFs
```{r}
# names for PDFs on disk
basename(pdf)
dirname(dirname(dirname(pdf)))
str_extract(pdf, "\\d{4}.pdf")

pdf_names <- str_extract(pdf, "\\d{4}.pdf")
```

```{r, eval=F}
# download pdfs
for(i in seq_along(pdf) ) {
  download.file(pdf[i], pdf_names[i], mode="wb")
}
```


# Transforming / Reading Data

## transforming PDFs - function

```{r}
# WINDOWS: xpdf: http://www.foolabs.com/xpdf/download.html 
#   add install path to path variable / see: http://www.computerhope.com/issues/ch000549.htm
# Linux: sudo apt-get install poppler-utils
pdftotext <- function(fname){
  fname_txt <- str_replace(fname, ".pdf", ".txt")
  system2(command = "pdftotext", args = fname)
  return(fname_txt)
}
```



## transforming PDFs - execution
```{r, eval=FALSE}
# transform PDFs to text
pdftotext(pdf_names[1])
pdftotext(pdf_names[2])
pdftotext(pdf_names[3])
pdftotext(pdf_names[4])
``` 



## loading text
```{r, warning=FALSE}
# laod text of PDF
text1 <- readLines("2013.txt", warn=FALSE)
```


## first glance at text
```{r}
substring(text1, 1, 60)[6:14]
```

- some useless/wrong characters &#8594; cleansing
- get rid of spaces
- get rid of parantheses
- information scheme is: \newline{} `FirstName Lastname, Institution (NumberOfReviews)`
- followed by actual extraction


##  preparation

```{r}
text1_tmp <- 
  text1 %>% 
  str_replace_all("[!\f]"," ") %>%   # drop form feed
  str_replace_all("\\]"," ") %>%     # drop ]
  str_replace_all("\\(|\\)", "") %>% # drop ( )
  str_replace(" ,", ",") %>%         # correct space
  str_replace_all("  ", " ") %>%     # correct space
  str_trim()                         # correct space

text1_tmp <- 
  text1_tmp[text1_tmp != ""] # drop empty lines

text1_tmp <- 
  text1_tmp[-c(1:5 )]        # drop non data
```

## cleaned up

```{r}
text1_tmp[1:10]
```



# First Result

## Reviewers

```{r}
length(grep("Konstanz", text1_tmp))
length(grep("Harvard", text1_tmp))
length(grep("Berlin", text1_tmp))
length(grep("Bamberg", text1_tmp))
length(grep("UCLA", text1_tmp))
```



# Extraction
  
## names

```{r}
names <- 
  text1_tmp %>% 
  str_extract("^.*?,") %>% 
  str_replace_all("  |,", " ") %>% 
  str_trim( )
sample(names, 10)
```


## institutions

```{r}
institution <-
  text1_tmp %>% 
  str_extract(",.*\\d") %>% 
  str_replace_all("^ ,|^,|\\d$","") %>% 
  str_trim()
sample(institution, 7)
```


## reviews

```{r}
reviews <-      
  text1_tmp %>% 
  str_extract("\\d+") %>% 
  as.numeric 
table(reviews)
```


## reviews

```{r}
data.frame(n=reviews, names, institution)[reviews > 3, ]
```


## save data gathered so far

```{r}
revdat <- data.frame(
  reviews, 
  names, 
  institution, 
  stringsAsFactors = FALSE
)
save(revdat, file = "revdat.Rdata")
```




# Extension (1) 

## geocoding institutions
```{r, message=FALSE}
require(ggmap)
# geocoding takes a while -> save results
# 2500 requests allowed per day
if ( !file.exists("scenario1_inst_geocoded_pos.rdata")){
  pos <- geocode(institution)
  geocodeQueryCheck() 
  save(pos, file="scenario1_inst_geocoded_pos.rdata")
} else {
  load("scenario1_inst_geocoded_pos.rdata")
}
```

## plot coordinates
```{r}
mapWorld <- borders("world")
map <-   
  ggplot() + 
  mapWorld + 
  geom_point(aes(x=pos$lon, y=pos$lat) ,
             color="#F54B1A90", size=3 ,
             na.rm=T) + 
  theme_bw() + 
  coord_map(xlim=c(-180, 180), ylim=c(-60,70))
```


## plot coordinates

```{r}
map # ajps 2013 reviewers worldwide
```



## Extension (2)

- grab articel authors for some years
- and compare to reviewers

















