---
title: "Web Data Collection with R"
author: "Peter Meißner / 2016-02-29 --  2016-03-04 / ECPR WSMT"
header-includes:
  - \definecolor{links}{HTML}{800080}
  - \hypersetup{colorlinks,linkcolor=,urlcolor=links}
output:
  beamer_presentation:
    fonttheme: structurebold
    keep_tex: yes
    toc: yes
  slidy_presentation: default
subtitle: HTML Forms POST Case Study
keep_tex: yes
---




# Posting HTML Forms




## Teaser
- ADCR author blog post on http://www.r-datacollection.com/ battle against each other 
 for the readability price 

## a first glance
```{r}
# ADCR: page 236
attr_inspector <- function(parsed_html, xpath){
  x <- html_nodes(parsed_html, xpath=xpath)
  x <- html_attrs(x)
  x <- lapply(x, function(x) as.data.frame(t(x)) )
  do.call(plyr::rbind.fill, x)
}
```

```{r}
library(httr)
library(rvest)
library(stringr)
```

## a first glance

```{r, cache=TRUE}
url  <- "http://read-able.com/"
html <- read_html(url)
attr_inspector( html, "//form")
attr_inspector( html, "//form[2]//input|//textarea|//select")
```


## HTTP messages
- but where goes our data? 

![](fig/httprequest.pdf)

## HTTP messages
- but where goes our data? 

![](fig/httpresponse.pdf)


## getting the texts
```{r, cache=TRUE}
dominic <- read_html("http://www.r-datacollection.com/blog/Fifty-years-of-Christmas-at-the-Windsors/")
dominic <- html_nodes(dominic, xpath="//p")
dominic <- html_text(dominic)
dominic <- str_c(dominic, collapse="\n")
```

## getting the texts
```{r, cache=TRUE}
peter <- read_html("http://www.r-datacollection.com/blog/Introduction-to-Public-Attention-Analytics-with-Wikipediatrend/")
peter <- html_nodes(peter, xpath="//p")
peter <- html_text(peter)
peter <- str_c(peter, collapse="\n")
```

## getting the texts
```{r}
simon <- read_html("http://www.r-datacollection.com/blog/Programming-a-Twitter-bot/")
simon <- html_nodes(simon, xpath="//p")
simon <- html_text(simon)
simon <- str_c(simon, collapse="\n")
```

## getting the texts
```{r}
christian <- read_html("http://www.r-datacollection.com/blog/Hassle-free-data-from-HTML-tables-with-the-htmltable-package/")
christian <- html_nodes(christian, xpath="//p")
christian <- html_text(christian)
christian <- str_c(christian, collapse="\n")
```


## posting texts

```{r}
force <- F # redo or not
if ( !file.exists("dominic.html") | force==T){
  resp_d <- POST("http://read-able.com/check.php",
            body=list(directInput=dominic), 
            encode="form")
  writeBin(content(resp_d, "raw"), 
           con="dominic.html" , useBytes=T)
}
```

## posting texts
```{r}
if ( !file.exists("peter.html")  | force==T){
  resp_p <- POST("http://read-able.com/check.php",
            body=list(directInput=peter), 
            encode="form")
  writeBin(content(resp_p, "raw"), 
           con="peter.html" , useBytes=T)
}
```

## posting texts
```{r}
if ( !file.exists("simon.html") | force==T ){
  resp_s <- POST("http://read-able.com/check.php",
            body=list(directInput=simon), 
            encode="form")
  writeBin(content(resp_s, "raw"), 
           con="simon.html" , useBytes=T)
}
```

## posting texts
```{r}
if ( !file.exists("christian.html")  | force==T){
  resp_c <- POST("http://read-able.com/check.php",
            body=list(directInput=christian), 
            encode="form")
  writeBin(content(resp_c, "raw"), 
           con="christian.html" , useBytes=T)
}
```


## the verdict
```{r}
verdict <- function(file){
  read_html(file) %>% 
    html_table() %>% 
    magrittr::extract2(1) %>% 
    magrittr::extract(2,)
}
```

## the verdict
```{r}
verdict("simon.html")
verdict("dominic.html")
verdict("peter.html")
verdict("christian.html")
```





# GET and POST within Developer Tools

## GET and POST within Developer Tools

Libve Clicking







































